
## sample metadata

|Sample.ID|OldSample.ID|Treatment|Sex|TreatmentN|Parent.ID|
|---------|------------|---------|---|----------|---------|
|12M      |S12M        |Exposed  |M  |3         |EM05     |
|13M      |S13M        |Control  |M  |1         |CM04     |
|16F      |S16F        |Control  |F  |2         |CF05     |
|19F      |S19F        |Control  |F  |2         |CF08     |
|22F      |S22F        |Exposed  |F  |4         |EF02     |
|23M      |S23M        |Exposed  |M  |3         |EM04     |
|29F      |S29F        |Exposed  |F  |4         |EF07     |
|31M      |S31M        |Exposed  |M  |3         |EM06     |
|35F      |S35F        |Exposed  |F  |4         |EF08     |
|36F      |S36F        |Exposed  |F  |4         |EF05     |
|39F      |S39F        |Control  |F  |2         |CF06     |
|3F       |S3F         |Exposed  |F  |4         |EF06     |
|41F      |S41F        |Exposed  |F  |4         |EF03     |
|44F      |S44F        |Control  |F  |2         |CF03     |
|48M      |S48M        |Exposed  |M  |3         |EM03     |
|50F      |S50F        |Exposed  |F  |4         |EF01     |
|52F      |S52F        |Control  |F  |2         |CF07     |
|53F      |S53F        |Control  |F  |2         |CF02     |
|54F      |S54F        |Control  |F  |2         |CF01     |
|59M      |S59M        |Exposed  |M  |3         |EM01     |
|64M      |S64M        |Control  |M  |1         |CM05     |
|6M       |S6M         |Control  |M  |1         |CM02     |
|76F      |S76F        |Control  |F  |2         |CF04     |
|77F      |S77F        |Exposed  |F  |4         |EF04     |
|7M       |S7M         |Control  |M  |1         |CM01     |
|9M       |S9M         |Exposed  |M  |3         |EM02     |


Getting sorted bams from bismark..


```{bash}
/opt/homebrew/bin/wget -r \
--no-check-certificate \
--quiet \
--no-directories --no-parent \
-P ../bg_data/ \
-A *sorted.bam',', \
https://gannet.fish.washington.edu/seashell/bu-mox/scrubbed/071521-cvBS/
```

https://gannet.fish.washington.edu/seashell/bu-mox/scrubbed/071521-cvBS/77F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',',



```{r include=FALSE}
file.list=list('../bg_data/12M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/13M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/16F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/19F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/22F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/23M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/29F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/31M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/36F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/39F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/3F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/41F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/44F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/48M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/50F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/52F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/53F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/54F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/59M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/64M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/6M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/76F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/77F_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/7M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam',
                '../bg_data/9M_R1_val_1_bismark_bt2_pe.deduplicated.sorted.bam'
)
```




```{r eval=FALSE, include=FALSE}
myobj_01 = processBismarkAln(location = file.list, sample.id = list("1","2","3","4","5","6","7","8"), assembly = "cv", read.context="CpG", mincov=2, treatment = c(0,0,0,0,1,1,1,1))
```

```{r include=FALSE}
#save(myobj_21, file = "../analyses/myobj_21")

```

```{r include=FALSE}
load("../analyses/myobj_21")
```
