---
title: "Identification of differentially expressed transcripts in C.virginica gonad exposed to elevated pCO2 using Ballgown"
author: "Sam White"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Use [Ballgown](https://github.com/alyssafrazee/ballgown) for identification of differentially expressed isoforms in _C.virginica_ gonad tissue exposed to elevated pCO<sub>2</sub>.

REQUIRES Linux-based system to run all chunks properly; some chunks will not work on Mac OS!

REQUIRES the following Bash programs:

- `wget`

- `tree`

- `md5sum`

REQUIRES the following R libraries:

- `[Ballgown](https://github.com/alyssafrazee/ballgown)` (Bioconductor)

- `tidyverse`

## Load `R` libraries

```{r}
library("ballgown")
library("tidyverse")
library("ggplot2")
```


## Set user variables!
```{r}
# Set maximum pvalue for isoform expression cutoff
pvalue <- 0.05

# Set maximum qvalue (false discovery rate) for isoform expression cutoff
qvalue <- 0.05
```



## Download Ballgown input files.

Notebooks detailing their creation:

- [FastQ trimming](https://robertslab.github.io/sams-notebook/2021/07/14/Trimming-C.virginica-Gonad-RNAseq-with-FastP-on-Mox.html)

- [Genome indexing, and exon/splice sites with HISAT2](https://robertslab.github.io/sams-notebook/2021/07/20/Genome-Annotations-Splice-Site-and-Exon-Extractions-for-C.virginica-GCF_002022765.2-Genome-Using-Hisat2-on-Mox.html)

- [Mapping and identificaion of isoforms with StingTie](https://robertslab.github.io/sams-notebook/2021/07/20/RNAseq-Alignments-C.virginica-Gonad-Data-to-GCF_002022765.2-Genome-Using-StringTie-on-Mox.html)

```{bash}
# Download Ballgown input files and directory structure
wget \
--directory-prefix ../data/ballgown \
--recursive \
--no-check-certificate \
--continue \
--cut-dirs 2 \
--no-host-directories \
--no-parent \
--quiet \
--reject "input_fastqs_checksums.md5" \
--accept "*.ctab,*checksums.md5" https://gannet.fish.washington.edu/Atumefaciens/20210726_cvir_stringtie_GCF_002022765.2_isoforms/
```

## Verify checksums

NOTE: Warnings are expected, as the checksums files have checksums for files that are not downloaded for this project.
```{bash}
cd ../data/ballgown

# Make a line
line="-----------------------------------------------------------------------------------------------"

# Set working directory
wd=$(pwd)

# Loop through directories and verify checksums
for directory in */
do
  cd "${directory}"
  # Get sample name; strips trailing slash from directory name
  sample="${directory%/}"
  
  echo ${line}
  echo "${sample}"
  echo ""
  
  # Confirm checksums; sorts for easier reading
  md5sum --check "${sample}"_checksums.md5 | sort -V
  echo ""
  echo "${line}"
  echo ""
  cd ${wd}
done

# Show downloaded directories/files
tree
```

## Find Ballgown installation location
```{r}
data_directory <-  system.file('extdata', package='ballgown') # automatically finds ballgown's installation directory
# examine data_directory:
data_directory
```

## Create Ballgown object
```{r}
# Uses regular expression in samplePattern to find all pertinent folders
# Load all measurement data
bg <-  ballgown(dataDir="../data/ballgown/", samplePattern='S.*[FM]', meas='all')
bg
```


## Download and filter metadata file

Filtered metadata will be used to create a phenotype dataframe needed for Ballgown differential expression analysis.

`TreatmentN` column explanation:

control males = 1
control females = 2
exposed males = 3
exposed females = 4


```{r create-dataframes-for-ballgwon-pData}
# Read in metadata file from URL
sample_metadata_full <- read.csv("https://raw.githubusercontent.com/epigeneticstoocean/2018_L18-adult-methylation/main/data/adult-meta.csv")

# View full metadata
head(sample_metadata_full)

# Subset metadata in preparation of creating pData data frame far Ballgown
# Sort by OldSample.ID to ensure matches directory structure (required for Ballgown)
sample_metadata_subset <- sample_metadata_full %>% select(OldSample.ID, Treatment, TreatmentN, Sex) %>% arrange(OldSample.ID)

# View subsetted metadata
head(sample_metadata_subset)

# Create modified metadata for testing differential expression processing
mutated_sample_metadata_subset <- sample_metadata_subset
mutated_sample_metadata_subset$TreatmentN[mutated_sample_metadata_subset$TreatmentN == 4] <- 0

head(mutated_sample_metadata_subset)

```
## Load phenotype dataframe into Ballgown object
```{r}
# Load phenotype info into Ballgown
pData(bg) <- sample_metadata_subset

# Examine phenotype data as it exists in Ballgown
phenotype_table <-  pData(bg)
head(phenotype_table)
```


## Look at exon info
```{r}
# Exon info
structure(bg)$exon
```

## Look at intron info
```{r}
# Intron info
structure(bg)$intron
```

## Look at transcript (isoform) info
```{r}
# Transcript info
structure(bg)$trans
```

## Load all transcript expression data
```{r}
# Expression data
whole_tx_table <-  texpr(bg, 'all')
whole_tx_table
```

## Load all gene expression data
```{r load-gene-expression-data}
whole_gx_table <-  gexpr(bg)
head(whole_gx_table)
```

## Generate boxplots to compare FPKM across all samples
```{r boxplots-to-compare-FPKM-across-all-samples} 
# Pull all transcript expression values stored in FPKM measurement from ballgown object
fpkm <- texpr(bg, meas = "FPKM")


# Log transform data and add 1 to all values to artificially prevent log2(0).
fpkm_df <- as.data.frame(log2(fpkm+1))
head(fpkm_df)

# Rotate data frame
# Creates a "Sex" column by pulling last character from each library name (M/F)
fpkm_df_pivot <- pivot_longer(
  fpkm_df, cols = starts_with("FPKM"), 
  names_to = "library") %>% 
  mutate(sex = (str_sub(library, -1)
                )
         )
head(fpkm_df_pivot)


# Sort data frame by sex
fpkm_df_pivot_sorted <- fpkm_df_pivot %>% arrange(sex)
head(fpkm_df_pivot_sorted)

# Set unique library names as vector
# Will be used to group boxplot by sex
fpkm_libraries_sorted_unique <- (unique(fpkm_df_pivot_sorted$library))
head(fpkm_libraries_sorted_unique)

# Re-rder data frame by sex-sorted data
fpkm_df_pivot$library <- factor(fpkm_df_pivot$library, levels = fpkm_libraries_sorted_unique)


# Produce boxplots of FPKM for each library
# Grouped by sex
ggplot(fpkm_df_pivot, aes(library, value, color = sex)) +
  geom_boxplot() +
  ggtitle("Comparisons of transcript FPKM values across all libraries") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("FPKM") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

# Save boxplot
ggsave(filename = "../figures/fpkm_f-vs-m_boxplot.pdf")

```

## Identify differentially expressed isofroms between Sex.
## Returns FoldChange, too (getFC = TRUE)
```{r}
# Set string describing comparison
# Used for final printout in chunk
comparison <- "female vs. male"


DET_sex_stat_results = stattest(bg, feature='transcript', meas='FPKM', covariate="Sex", getFC = TRUE)
head(DET_sex_stat_results)

# Filter based on p-value and q-value
DET_sex_filtered_stat_results <- filter(DET_sex_stat_results, pval <= pvalue & qval <= qvalue)
head(DET_sex_filtered_stat_results)

# Merge with full table to get subset of just differentially expressed isoforms
merged_DET_sex_filtered <- merge(x = DET_sex_filtered_stat_results, y = whole_tx_table, by.x = "id", by.y = "t_id")
head(merged_DET_sex_filtered)

# Convert to BED coordinate system, which is 0-based start
# and end EXCLUSIVE; so need to subtract 1 from start coordinates ONLY
merged_DET_sex_filtered <- merged_DET_sex_filtered %>% mutate(start = start - 1)
head(merged_DET_sex_filtered)

# Filter for female up-regulated transcripts (i.e. fold-change < 1)
# Convert to BED coordinate system, which is 0-based start
# and end EXCLUSIVE; so need to subtract 1 from start coordinates ONLY
DET_sex_filtered_female <- filter(merged_DET_sex_filtered, fc < 1) %>% mutate(start = start - 1)
head(DET_sex_filtered_female)

# Filter for male up-regulated transcripts (i.e. fold-change < 1)
# Convert to BED coordinate system, which is 0-based start
# and end EXCLUSIVE; so need to subtract 1 from start coordinates ONLY
DET_sex_filtered_male <- filter(merged_DET_sex_filtered, fc > 1) %>% mutate(start = start - 1)
head(DET_sex_filtered_male)

# Count number of DET.
count_DET_sex_filtered_stat_results <-  nrow(DET_sex_filtered_stat_results)

# Print number of DET
cat("Number of", comparison, "DET with p-values and q-values <= ", pvalue, ":", count_DET_sex_filtered_stat_results, "\n")

# Count number of female DET.
count_DET_sex_filtered_female <-  nrow(DET_sex_filtered_female)

# Print number of female DET
cat("Number of female DET with p-values and q-values <= ", pvalue, ":", count_DET_sex_filtered_female, "\n")

# Count number of male DET.
count_DET_sex_filtered_male <-  nrow(DET_sex_filtered_male)

# Print number of female DET
cat("Number of male DET with p-values and q-values <= ", pvalue, ":", count_DET_sex_filtered_male, "\n")

# Write merged dataframe to CSV
write.csv(
  merged_DET_sex_filtered, 
  file = "../analyses/merged_DET_sex_filtered_p-0.05_q-0.05.csv", 
  quote = FALSE, 
  row.names = FALSE)


# Creates a BED file of all DETs and inserts necessary columns to create properly formatted BED file. 
write.table(
  (
    merged_DET_sex_filtered %>% 
    select(chr, start, end, t_name, strand) %>% 
      add_column(
        score = "0", # Inserts a "score" column and assigns a value of "0" to all rows.
        .before = "strand"
        )
    ), 
  file = "../analyses/DET_sex_filtered_p-0.05_q-0.05.bed",
  sep = "\t",
  col.names = FALSE,
  row.names = FALSE,
  quote = FALSE
)

# Creates a BED file of only female DETs and inserts necessary columns to create properly formatted BED file. 
write.table(
  (
    DET_sex_filtered_female %>% 
    select(chr, start, end, t_name, strand) %>% 
      add_column(
        score = "0", # Inserts a "score" column and assigns a value of "0" to all rows.
        .before = "strand"
        )
    ), 
  file = "../analyses/DET_sex_female_filtered_p-0.05_q-0.05.bed",
  sep = "\t",
  col.names = FALSE,
  row.names = FALSE,
  quote = FALSE
)

# Creates a BED file of only male DETs and inserts necessary columns to create properly formatted BED file. 
write.table(
  (
    DET_sex_filtered_male %>% 
    select(chr, start, end, t_name, strand) %>% 
      add_column(
        score = "0", # Inserts a "score" column and assigns a value of "0" to all rows.
        .before = "strand"
        )
    ), 
  file = "../analyses/DET_sex_male_filtered_p-0.05_q-0.05.bed",
  sep = "\t",
  col.names = FALSE,
  row.names = FALSE,
  quote = FALSE
)
```

## Identify differentially expressed genes (DEG) between Sex.
## Returns FoldChange, too (getFC = TRUE)
```{r female-male-DEG}

# Set string describing comparison
# Used for final printout in chunk
comparison <- "female vs. male"

# Determine differentially expressed genes with ballgown
DEG_sex_stat_results = stattest(bg, feature='gene', meas='FPKM', covariate="Sex", getFC = TRUE)
head(DEG_sex_stat_results)

# Filter based on p-value and q-value
DEG_sex_filtered_stat_results <- filter(DEG_sex_stat_results, pval <= pvalue & qval <= qvalue)
head(DEG_sex_filtered_stat_results)


# Merge with full table to get subset of just differentially expressed isoforms
merged_DEG_sex_filtered <- merge(x = DEG_sex_filtered_stat_results, y = whole_tx_table, by.x = "id", by.y = "gene_id")
head(merged_DEG_sex_filtered)


# Write merged dataframe to CSV
write.csv(
  merged_DEG_sex_filtered, 
  file = "../analyses/merged_DEG_sex_filtered_p-0.05_q-0.05.csv", 
  quote = FALSE, 
  row.names = FALSE)

# Inspect full merged file
head(merged_DEG_sex_filtered)

# Convert to BED coordinate system, which is 0-based start
# and end EXCLUSIVE; so need to subtract 1 from start coordinates ONLY
merged_DEG_sex_filtered <- merged_DEG_sex_filtered %>% mutate(start = start -1)
head(merged_DEG_sex_filtered)

# Creates a BED file and inserts necessary columns to create properly formatted BED file. 
write.table(
  (
    merged_DEG_sex_filtered %>% 
    select(chr, start, end, t_name, strand) %>% 
      add_column(
        score = "0", # Inserts a "score" columns and assigns a value of "0" to all rows.
        .before = "strand"
        )
    ), 
  file = "../analyses/DEG_sex_filtered_p-0.05_q-0.05.bed",
  sep = "\t",
  col.names = FALSE,
  row.names = FALSE,
  quote = FALSE
)

# Count number of DET.
# Subtract "1" to account for header row.
count_DEG_sex_filtered_stat_results <-  nrow(DEG_sex_filtered_stat_results)

# Print number of DEG
cat("Number of ", comparison, " DEG with p-values and q-values <= ", pvalue, ":", count_DEG_sex_filtered_stat_results)
```


## Identify differentially expressed isofroms between Treatment.
## NOTE: Will not return fold change; not possible in multi-group comparisons
```{r}
stat_results = stattest(bg, feature='transcript', meas='FPKM', covariate="TreatmentN")
head(stat_results)

# Filter DET based on set p-avlues and q-values
filtered_stat_results <- filter(stat_results, pval <= pvalue & qval <= qvalue)
head(filtered_stat_results)

# Count number of DET.
# Subtract "1" to account for header row.
count_filtered_stat_results <-  nrow(filtered_stat_results) - 1

# Print number of DET
cat("Number of DET with p-values and q-values <= ", pvalue, ":", count_filtered_stat_results)
```


## Examine if differential expression results are affected by TreatmentN numbering
## I.e. are DET based off of alphanumeric comparisons?
## Change one of the `TreatmentN` numbers to `0` and compare to results from analysis in chunk above.
```{r}
# Load phenotype info into Ballgown
pData(bg) <- mutated_sample_metadata_subset
# Examine phenotype data as it exists in Ballgown
phenotype_table <-  pData(bg)

stat_results = stattest(bg, feature='transcript', meas='FPKM', covariate="TreatmentN")
head(stat_results)

# Filter DET based on set p-avlues and q-values
filtered_stat_results <- filter(stat_results, pval <= pvalue & qval <= qvalue)
head(filtered_stat_results)

# Count number of DET.
# Subtract "1" to account for header row.
count_filtered_stat_results <-  nrow(filtered_stat_results) - 1

# Print number of DET
cat("Number of DET with p-values and q-values <= ", pvalue, ":", count_filtered_stat_results)
```


## Identify differentially expressed transcripts between Treatment (control/exposed).
## Returns FoldChange, too (getFC = TRUE)
```{r control-exposed-DET}
# Load phenotype info into Ballgown
pData(bg) <- sample_metadata_subset
# Examine phenotype data as it exists in Ballgown
phenotype_table <-  pData(bg)

stat_results = stattest(bg, feature='transcript', meas='FPKM', covariate="Treatment", getFC = TRUE)
head(stat_results)

# Filter DET based on set p-avlues and q-values
filtered_stat_results <- filter(stat_results, pval <= pvalue & qval <= qvalue)
head(filtered_stat_results)

# Count number of DET.
# Subtract "1" to account for header row.
count_filtered_stat_results <-  nrow(filtered_stat_results) - 1

# Print number of DET
cat("Number of DET with p-values and q-values <= ", pvalue, ":", count_filtered_stat_results)
```
